Simulations on Adaptive Kernel Smoothers

### 2020-11-03 src/Simulations_ver1.Rmd

---
title: Simulations on Adaptive Kernel Smoothers
author: Qiuyi Wu, Cuong Pham, Xing Qiu
date: "`r format(Sys.time(), '%B %d, %Y')`"
abstract: "Optional: An abstract with font Arial 10 to describe the study."
output:
  pdf_document:
    fig_height: 6
    fig_width: 6
    fig_caption: true
---

@@

set.seed(123)
for (j in 1:reps){
  print(paste0("Repetition: ", j))
  y <- u + sd1*rnorm(N)
  yhat.adapt <- ks(x, y, bandwidth=b)
  yhat.fixed <- sapply(logbs, function(lb) ks(x,y,bandwidth=10^lb))
  Res <- sweep(cbind(yhat.adapt,yhat.fixed), 1, u)
  MSEtab[j,] <- colMeans(Res^2)
}

@@

## Sim2: Estimating a Periodic Function on $S^1$

More work is needed.

>>>>

---
title: "Simulations on Adaptive Kernel Smoothers"
author: "Qiuyi Wu, Cuong Pham, Xing Qiu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
  html_document:
    df_print: paged
abstract: 'Optional: An abstract with font Arial 10 to describe the study.'
---

@@

set.seed(123)
for (j in 1:reps){
  print(paste0("Repetition: ", j))
  y <- u + sd1*rnorm(N)
  yhat.adapt <- ks(x, y, bandwidth=b, type = "Euclidean")
  yhat.fixed <- sapply(logbs, function(lb) ks(x,y,bandwidth=10^lb, type = "Euclidean"))
  Res <- sweep(cbind(yhat.adapt,yhat.fixed), 1, u)
  MSEtab[j,] <- colMeans(Res^2)
}

@@

## Sim2: Estimating a Periodic Function on $[-2\pi, 2\pi]$
```{r sim2-true-signal, fig.cap='Optimal bandwidth computed based on our theoretical derivations.'}

## specify parameters that generates data in Sim2
N <- 501
## Note that you shouldn't use "sd" as an object; it *replaces* an
## important system function sd()!!
sd1 <- 1; var1 <- sd1^2

## an evenly-spaced grid of x defined on [-5,5]
x <- seq(-2*pi, 2*pi, length.out=N); dx <- 10/(N-1)
## the true function to be estimated
u <- sin(x)
## its second order derivative
D2u <- cos(x)
## the optimal bandwidth
b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)

## the fixed bandwidths
min.logb <- -3; max.logb <- 3; nb <- 11
logbs <- seq(min.logb, max.logb, length.out=nb)

```
In this simulation, we aim to estimate the following function
$$u(x) := \sin(x), \qquad x\in [-2\pi, 2\pi] $$

Again we use the same grid (a total number of N=`r N` evenly spaced points) on $[-2\pi, 2\pi]$. The observed data are the summation of $u(x_{i})$ and $i.i.d.$ normal noise $\epsilon_{j} \sim N(0, \sigma^2)$, $\sigma = `r sd1`$.

```{r Sim2, fig.cap="Relationship between the smoothing bandwidth and mean MSE. The thick horizontal broken line on the bottom represents the mean MSE of the optimal adaptive kernel smoother; dots and thin solid curve represent the mean MSE obtained from kernel smoother with various fixed bandwidths; dotted curves represent +/- two standard errors from the mean MSEs. "}
MSEtab <- matrix(0, nrow=reps, ncol=nb+1)
colnames(MSEtab) <- c("Adaptive", paste0("logb=",logbs))
set.seed(123)
for (j in 1:reps){
  print(paste0("Repetition: ", j))
  y <- u + sd1*rnorm(N)
  yhat.adapt <- ks(x, y, bandwidth=b, type = "circular")
  yhat.fixed <- sapply(logbs, function(lb) ks(x,y,bandwidth=10^lb, type = "circular"))
  Res <- sweep(cbind(yhat.adapt,yhat.fixed), 1, u)
  MSEtab[j,] <- colMeans(Res^2)
}

MSE.mean <- colMeans(MSEtab); MSE.se <- apply(MSEtab, 2, sd)/sqrt(reps)
plot(logbs, MSE.mean[-1], type="b", ylim=range(MSE.mean),
     xlab="log-bandwidth", ylab="Mean MSE")
## +/- 2*STDERR
lines(logbs,MSE.mean[-1]+2*MSE.se[-1], lty=3)
lines(logbs,MSE.mean[-1]-2*MSE.se[-1], lty=3)
abline(h=MSE.mean[1], lwd=2, lty=2)
```

From this figure, we see that the adaptive smoother cannot beat the fixed bandwidth smoother. Fixed bandwidth smoother with `r names(which(MSE.mean==min(MSE.mean)))` achieved the smallest mean MSE.

```{r Sim2.smoothing.plot, fig.cap="The true function and functions estimated by various smoothing methods. True underlying function is marked in black, adaptive smoothing in red, two choices of fixed bandwidth parameters are in green (small bandwidth, $\\log_{10} b=-0.6$) and blue (the smallest fixed bandwidth, $\\log_{10} b=0$). The left panel shows all data points, and the right panel is a zoomed-in version which focuses on the right boundary. ", fig.width=12, fig.height=6}

plot(x, y, pch="."); lines(x,u)
lines(x, yhat.adapt, col="red")
lines(x, yhat.fixed[, 5], col="green")
lines(x, yhat.fixed[, 6], col="blue")
